---
title: "cmei24_OriginalHomeworkCode_03"
author: "Christian Mei"
format: html
editor: visual
---

# **Some of my best friends are Zombies…**

```{r}
library(tidyverse)
library(curl)
```

Load the zombie dataset **from** GitHub

```{r}
zombie_data <- curl(url = 'https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/refs/heads/master/AN588_Spring25/zombies.csv')

zombie_data <- read_csv(zombie_data, col_names = TRUE)

str(zombie_data)
```

```{r}
head(zombie_data)
```

## Calculate the *population* mean and standard deviation for each quantitative random variable

```{r}
# Population variance function
pop_v <- function(x) {
    sum((x - mean(x))^2)/(length(x))
}

# Population sd function (square root of variance)
pop_sd <- function(x) {
    sqrt(pop_v(x))
}
```

Use for loop to pick which columns we want to use

```{r}

for (column in colnames(zombie_data)){
  col_values <- zombie_data[[column]] # extract the data from each column picked
  if (column != "id" && is.numeric(col_values)){ # We only want values that are numeric but we also don't want to include ID
    average <- mean(col_values)
  variance <- pop_v(col_values)
  s_dev <- pop_sd(col_values)
  print(paste(column, " has a mean of ", average, " a population variance of ", variance, " and a standard deviation of ", s_dev))
  print("------------------------------------------------------------------------------")
  }
}
```

## Use {ggplot} to make boxplots of each of these variables by gender.

```{r}
for (column in colnames(zombie_data)){
  col_values <- zombie_data[[column]]
  if (column != "id" && is.numeric(col_values)){
    print(
      ggplot(zombie_data, aes(x = as.factor(gender), y = col_values, fill = as.factor(gender))) +
  geom_boxplot() + 
  labs(y = column, x = "Gender", fill = "Gender")
    ) +
      theme_classic()
  }
}

```

## Use {ggplot} to make scatterplots of height and weight in relation to age. Do these variables seem to be related? In what way?

Height vs. Age

```{r}
ggplot(zombie_data, aes(x = age, y = height)) + 
  geom_point(color = "red") +
  geom_smooth(method = "lm", color = "black", se = TRUE) +
  labs(x = "Age", y = "Height") + 
  theme_minimal()
```

```{r}
cor(x = zombie_data$age, y = zombie_data$height, method = "spearman")
```

Weight vs. Age

```{r}
ggplot(zombie_data, aes(x = age, y = weight)) + 
  geom_point(color = "darkred") +
  geom_smooth(method = "lm", color = "black", se = TRUE) +
  labs(x = "Age", y = "Weight") + 
  theme_minimal()
```

```{r}
cor(x = zombie_data$age, y = zombie_data$weight, method = "spearman")
```

## Using histograms and Q-Q plots, check whether the quantitative variables seem to be drawn from a normal distribution. Which seem to be and which do not

```{r}

par(mfrow = c(2, 3)) # Using this to show multiple figures c(rows, columns)
for (column in colnames(zombie_data)){
  col_values <- zombie_data[[column]]
  if (column != "id" && is.numeric(col_values)){
    qqnorm(col_values, main = paste("Normal QQ plot ", column))
    qqline(col_values, col = "red")
  }
}

par(mfrow = c(1, 1)) # Reset it so the next figures don't have the same configuration
```

```{r}

par(mfrow = c(2, 3))
for (column in colnames(zombie_data)){
  col_values <- zombie_data[[column]]
  if (column != "id" && is.numeric(col_values)){
    print(
      ggplot(zombie_data, aes(x = col_values)) +
        geom_histogram(color = "black", fill = "lightblue") + 
        labs(y = "Frequency", x = column)) + 
        theme_minimal()
  }
}



```

Discrete (zombies_killed, and years_of_education) values rather than continuous

## Now use the sample() function to sample ONE subset of 30 zombie survivors (without replacement) from this population and calculate the mean and sample standard deviation for each variable.

```{r}
set.seed(1)
sample_rows <- sample(nrow(zombie_data), 30)
zombie_data_sample <- zombie_data[sample_rows,]
zombie_data_sample
```

```{r}
# Standard Error
SE <- function(x) {
    sqrt(var(x)/length(x))
}

# Confidnce Interval
normalCI = function(x, CIlevel = 0.95) {
    upper = mean(x) + qnorm(1 - (1 - CIlevel)/2) * sqrt(var(x)/length(x))
    lower = mean(x) + qnorm((1 - CIlevel)/2) * sqrt(var(x)/length(x))
    ci <- c(lower, upper)
    return(ci)
}
```

```{r}
for (column in colnames(zombie_data_sample)){
  col_values <- zombie_data_sample[[column]] # extract the data from each column picked
  if (column != "id" && is.numeric(col_values)){ # We only want values that are numeric but we also don't want to include ID
    average <- mean(col_values)
    variance <- var(col_values) # we can use var() now because its a sample
    s_dev <- sd(col_values) # we can use sd() now because its a sample
    s_error <- SE(col_values)
    print(paste(column, " has a mean of ", average, " a population variance of ", variance, " a standard deviation of ", s_dev, " and a standard error of ", s_error))
    if(column != 'zombies_killed' && column != 'years_of_education'){
      confidence_interval <- normalCI(col_values)
      print(paste("Lower lvl confidence intervals for", column, "is:", confidence_interval[1]))
      print(paste("Upper lvl confidence intervals for", column, "is:", confidence_interval[2]))
    }
    print("------------------------------------------------------------------------------")
  }
}
```

Bootstrapping

```{r}
zombie_means <- mean(zombie_data_sample)
n <- 30
for (i in 2:99) {
    zombie_means[i] <- mean(sample(zombie_data, n, replace = TRUE))
    # Cannot take a sample larger than the population when 'replace = FALSE'
}
```
